Parallel Hierarchical Clustering using \\ Rank-Two Nonnegative Matrix Factorization \thanks{Identify applicable funding agency here. If none, delete this.}

### 2020-07-16 latex/tex/experiments.tex

\Cref{fig:synrank2speedup,fig:rwrank2speedup} show performance relative to the smallest number of compute nodes required to store data and factor matrices.
The synthetic data is chosen to fit on a single node, while the \image{} data requires 10 compute nodes to store the input matrix as well as temporary and output matrices.
For these data sets, we observe nearly perfect strong scaling, with 47$\times$ speedup on 50 compute nodes (over 1 compute node) for synthetic data and 7.1$\times$ speedup on 80 compute nodes (over 10 compute nodes) for \image{} data.

@@

\GB{Do these breakdowns include the score function?  actually this may explain why all-reduce costs more, it also has $O(n)$ data for score...}

\GB{4 plots: 2 data sets with speedup/breakdown, note that we include time for complete levels, ignoring nodes in incomplete levels for easier interpretation}

From equation \ref{eq:treecost}, we expect to see perfect strong scaling in a computationally bound hierarchical clustering problem with fixed frontier node $k=100$ count.
The problem should become latency bound with its $\log p$ cost as we scale.

@@

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/synthetic_hierarchical_speedup.pdf}
\caption{Strong Scaling Speedup for Hierarchical Clustering. The total time taken for complete Hierarchical Clustering \cref{alg:hiernmf2}}
\label{fig:synhierspeedup}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/synthetic_hier_strongscaling.pdf}
\caption{Relative Time Breakdown for Hierarchical Clustering. The plot uses the total time taken for complete Hierarchical Clustering \cref{alg:hiernmf2}}
\label{fig:synhierstrongscaling}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/realworld_hierarchical_speedup.pdf}
\caption{Strong Scaling Speedup for Hierarchical Clustering on \image{} Data. The total time taken for complete Hierarchical Clustering \cref{alg:hiernmf2}}
\label{fig:rwhierspeedup}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/realworld_hier_strongscaling.pdf}
\caption{Relative Time Breakdown for Hierarchical Clustering on \image{}. The plot uses the total time taken for complete Hierarchical Clustering \cref{alg:hiernmf2}}
\label{fig:rwhierstrongscaling}
\end{center}
\end{figure}

>>>>

\Cref{fig:synrank2speedup,fig:rwrank2speedup} show performance relative to the smallest number of compute nodes required to store data and factor matrices.
For these data sets, we observe nearly perfect strong scaling, with 42$\times$ speedup on 40 compute nodes (over 1 compute node) for synthetic data and 7.1$\times$ speedup on 80 compute nodes (over 10 compute nodes) for \image{} data.

@@

From \Cref{eq:treecost}, we expect to see perfect strong scaling in a computationally bound hierarchical clustering problem with target cluster count $k=100$.
As $k$ is large, we expect the latency cost of small problems deep in the tree to limit scalability.

@@

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/synthetic_hierarchical_speedup.pdf}
\caption{Strong Scaling Speedup for Hierarchical Clustering on Synthetic data}
\label{fig:synhierspeedup}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/synthetic_hier_strongscaling.pdf}
\caption{Relative Time Breakdown for Hierarchical Clustering on Synthetic data}
\label{fig:synhierstrongscaling}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/realworld_hierarchical_speedup.pdf}
\caption{Strong Scaling Speedup for Hierarchical Clustering on \image{} Data}
\label{fig:rwhierspeedup}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/realworld_hier_strongscaling.pdf}
\caption{Relative Time Breakdown for Hierarchical Clustering on \image{} Data}
\label{fig:rwhierstrongscaling}
\end{center}
\end{figure}
